### 2004-07-13 Documentation/part1.tex
Minor changes only The spaces in index entries with \_ remains, it seems to be rooted in the way how \index processes its arguments -> we have to edit the idn-files at last by hand :-<

\index{norm}{Experiment!list}
\index{norm}{Initialization!experiments}
For storing the experiments, \GLOBES\ uses the initially empty list of experiments \GLB{\_experiment\_list}. To add a pre-defined experiment to this list, one can use the function {\tt glbInitExperiment}:
\begin{function}
\index{norm}{Experiment initialization}
\GLBNS{InitExperiment}
{\tt int glbInitExperiment(char *inf, glb\_exp *in, int *counter)}
 adds a single experiment with the filename {\tt inf} to the list of currently loaded experiments. The {\tt counter} is a pointer to the 
 variable containing the number of experiments, and the experiment {\tt in}
 points to the beginning of the experiment list. The function returns
 zero if it was successful. 
\end{function}
Normally, a typical  call of {\tt glbInitExperiment} is 
\begin{quote}
{\tt glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0],\\  \hspace*{8cm} \&glb\_num\_of\_exps); }
\end{quote}
In this case, the experiment in the file {\tt NuFact.glb} is added to the internal global list of experiments, and the experiment counter is increased. The
experiment then has the number {\tt glb\_num\_of\_exps}-1. The elements
of the experiment list have the type \GLB{\_exp}, which the
user will not need to access directly in most cases. The experiment definition files, which usually end with {\tt .glb}, and any
supporting files, are first of all searched in the current directory, and then in the global path variable \GLBC{GLB\_PATH} set in your environment.
A list of pre-defined experiment prototypes, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}. If the program cannot find these files, or some of them are syntactically not correct, it will break at this place.

>>>>

\index{norm}{Experiment!list}
\index{norm}{Initialization!experiments}
For storing the experiments, \GLOBES\ uses the initially empty list of experiments \GLB{\_experiment\_list}. To add a pre-defined experiment to this list, one can use the function {\tt glbInitExperiment}:
\begin{function}
\index{norm}{Experiment initialization}
\GLBNS{InitExperiment}
{\tt int glbInitExperiment(char *inf, glb\_exp *in, int *counter)}
 adds a single experiment with the filename {\tt inf} to the list of currently loaded experiments. The {\tt counter} is a pointer to the 
 variable containing the number of experiments, and the experiment {\tt in}
 points to the beginning of the experiment list. The function returns
 zero if it was successful. 
\end{function}
Normally, a typical  call of {\tt glbInitExperiment} is 
\begin{quote}
{\tt glbInitExperiment("NuFact.glb",\&glb\_experiment\_list[0],\\  \hspace*{8cm} \&glb\_num\_of\_exps); }
\end{quote}
In this case, the experiment in the file {\tt NuFact.glb} is added to the internal global list of experiments, and the experiment counter is increased. The
experiment then has the number {\tt glb\_num\_of\_exps}-1. The elements
of the experiment list have the type \GLB{\_exp}, which the
user will not need to access directly in most cases. The experiment definition files, which usually end with {\tt .glb}, and any
supporting files, are first of all searched in the current directory, and then in the global path variable 
%
\index{norm}{GLB PATH@{\tt GLB\_PATH}}
\index{norm}{Environment variables!GLB PATH@{\tt GLB\_PATH}}
%
set in your environment.
\index{norm}{Path resolution}
A list of pre-defined experiment prototypes, their filenames, their short descriptions, and the references of their definitions can be found in \tabl{experiments}. If the program cannot find these files, or some of them are syntactically not correct, it will break at this place.

### 2004-07-13 Documentation/part2.tex
Minor changes only The spaces in index entries with \_ remains, it seems to be rooted in the way how \index processes its arguments -> we have to edit the idn-files at last by hand :-<

Eventually, the total event rate $x_i$ in a bin $i$ is given by
\begin{equation}
x_i(a,b,c,d)=s_i(a,b)+b_i(c,d) \, ,
\end{equation}
and is thus a function of four parameters. 
The four parameters $a,b,c,d$ have been introduced in order to describe
systematical uncertainties and are the nuisance parameters.
Each of the four parameters has a central value and systematical error.
The central values for all of
the four parameters have to be always defined. They are called 
signal normalization ($a$), signal tilt/calibration ($b$), 
background  normalization ($c$) and
background tilt/calibration ($d$). The default values are
\begin{equation}
a=1\,,\quad b=0\,,\quad c={\tt not \, \, assigned}\,,\quad d=0\,.
\end{equation}
Thus, for the background normalization $c$, the value has to be specified 
in \emph{either} case. The values for the normalization and the 
values of tilt/calibration are always regarded
as a pairs, \ie, they are given in the form {\tt normalization : tilt}. The errors are treated in the same way. For example, we have
\begin{quote}
{\tt
\tb @signalerror =       0.001  :       0.01\\
\tb @backgroundcenter =  0.1 :       0.0\\
\tb @backgrounderror =   0.001 :       0.01
}
\end{quote}
There is no {\tt @signalcenter} in this definition, 
since by default the central value for the
signal normalization is $1$ and the central value for the tilt/calibration 
is $0$.

The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
parameters which are minimized over. This choice is specified with the 
error dimension variable\index{norm}{Error dimension}, and the different 
possibilities are shown in
\Tab~\ref{tab:error_dim}. Since the error dimension defines the treatement
of systematics it is useful to have define a matched pair of error dimensions
for each rule, where on value describes how the event rate is computed
for \emph{no} systematics and the other one is with systematics 
(see also~\Sec~\ref{sec:systematics}).
The error dimensions for the case of no systematics is set for each rule 
by the value of {\tt @errordim\_sys\_off}\index{aedl}{{\tt rule}!{\tt "@errordim\_sys\_off}}, whereas the error dimension
for systematics on is given by {\tt @errordim\_sys\_on}
\index{aedl}{{\tt rule}!{\tt "@errordim\_sys\_on}}. Thus, for example,
the complete error dimension definition could look like  
\begin{quote}
{\tt 
\tb @errordim\_sys\_off = 2\\
\tb @errordim\_sys\_on = 0
}
\end{quote}
It is foreseen to add the possiblity to extend the set 
of error dimensions or the set of possible systematical errors in one
of the next versions of \GLOBES.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\begin{table}[t!]
\begin{center}
\begin{tabular}[h]{|c|cccc|c|l|}
\hline
Error dimension&$a$&$b$&$c$&$d$&Tilt/Calibration&Remarks\\
\hline
\hline
0&+&+&+&+&T&Systematics with tilt\\
2&-&-&-&-&-&No systematics\\
4&+&+&+&+&T&Total rates\\
7& $\infty$ &-& $\infty$ &-&-&Spectrum only\\
8&-&-&-&-&-&Total rates, No systematics\\
9&+&+&+&+&C&Systematics with calibration\\
\hline
\end{tabular}
\caption[Table of error dimensions]{\label{tab:error_dim}
Possible values of the error dimensions variable in \GLOBES\ and their meaning. If a parameter is designated with $+$, it will
be marginalized over, and therefore the corresponding error needs to 
have a non-zero value. If the cases with ``total rates'' in the remarks, 
the summation over the bins is performed \emph{before} computing 
the $\chi^2$, \ie, no spectral information is used. 
The error dimension~7 ``spectrum only'' leaves the 
normalization free ($\sigma_a=\sigma_c=\infty$), and therefore 
only the spectral information is used. As a consequence,
the settings for the normalization error will be ignored (designated 
with the symbol $\infty$). 
 }
\end{center} 
\end{table} 
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\AEDL\ is a powerful language to describe a vast variety 
of different experiments.
This chapter demonstrates how to test whether a given \AEDL\ file describes
a given experiment. To this end the \GLOBES\ package also features a program
called {\tt globes}\index{norm}{{\tt globes}}, which can be seen either as a 
crude \AEDL\ debugger 
or a simple command line oriented tool to convert the rather 
abstract \AEDL\ experiment description into more accessible event rates.

@@

The {\tt globes}\index{norm}{{\tt globes}} binary is installed along 
with the library \index{norm}{{\tt libglobes}}, but
into the directory {\tt \$prefix/bin/}. In order to use the {\tt globes} 
utility this directory\footnote{This is automatically the case, when
no options have been given to {\tt configure}, and {\tt make install} was 
executed with root-privilege, \ie\ a standard installation was done.} has to 
be in the path of the shell used to invoke it.

@@

\index{norm}{{\tt globes}!oscillation parameters}
Calling {\tt globes} without any options and with a {\tt .glb}-file as argument produces an event summary at rule level, taking into account 
the full description in the file, \ie\ all efficiencies, backgrounds and resolution effects. Thus, these are the event rates which are actually 
used to compute the $\chi^2$. By default the oscillation parameters used to
calculate the transition probability are
\begin{eqnarray}
\label{eq:globes_params}
\sin^22\theta_{12}=0.8&\quad&\Delta m^2_{21}=7\cdot10^{-5}
\,\mathrm{eV}^2\,,\nonumber\\
\sin^22\theta_{23}=1.0&\quad&\Delta m^2_{31}=3\cdot10^{-3}
\,\mathrm{eV}^2\,,\nonumber\\
\delta=0&\quad&\sin^22\theta_{13}=0.1\,.\
\end{eqnarray}
Of course, it is possible to change these default values either by using the
option {\tt -p} on a call by call basis or by setting the environment variable
{\tt GLB\_CENTRAL\_VALUES}\index{norm}{{\tt  GLB\_CENTRAL\_VALUES}}.
\begin{quote}
{\tt
globes -p'0.55,0,0.785,0,0.0008,0.0025'\\
globes \verb+--+parameters='0.55,0,0.785,0,0.0008,0.0025'\\
}
\end{quote}
or by defining within the shell session or in the shell profile:
\begin{quote}
{\tt
export GLB\_CENTRAL\_VALUES='0.55,0,0.785,0,0.0008,0.0025'
}
\end{quote}
Furthermore it is possible to switch off oscillations altogether with t
he {\tt -N} option
and to switch them on again with {\tt -O} (the default). The effect of {\tt -N}
is the same than to prefix all channels with {\tt NOSC\_}.

\index{norm}{{\tt globes}!warnings}\index{norm}{{\tt globes}!errors}
\index{norm}{{\tt globes}!verbosity}
The \AEDL\ parser and interpreter has basically three levels of messages to
th user: warnings, errors and fatal errors. Fatal errors always are reported
and the program exits with status '1'. Usually only errors and no warnings
are reported. The verbosity level can be chosen by the {\tt -v} option, where
{\tt -v1} is default, \ie\ only errors and fatal errors are reported. {-v0}
corresponds to reporting fatal errors only, whereas {\tt -v2} will
print in addition to fatal errors and errors also warnings. It is recommended
to test any new {\tt .glb}-file with {\tt -v2} to see the warnings at least 
once and to decide whether there is a problem and in case fix it. With {\tt -v3}
all files read by {\tt globes} are displayed together with their path and
with  {\tt -v4} all files which have been attempted to be read are shown.
These two setting are useful to clarify path resolution issues and shadowing
of file names.

@@

\index{norm}{{\tt globes}!total rates}
\index{norm}{{\tt globes}!spectral rates}
By default {\tt globes} returns total rates corresponding to the 
{\tt -t} option, this can be changed to
to a full spectrum by using {\tt -s}. The spectral rates are shown in a
table where the first columns always gives the central energy of the
corresponding bin.

@@

\subsection*{Channel level}
\index{norm}{{\tt globes}!channel rates}

@@

\subsection*{Rule level}
\index{norm}{{\tt globes}!rule rates}

@@

\subsection*{Output}
\index{norm}{{\tt globes}!output}

@@

\subsection*{\AEDL\ variable substitution}
\index{norm}{{\tt globes}!variable substitution}

>>>>

Eventually, the total event rate $x_i$ in a bin $i$ is given by
\begin{equation}
x_i(a,b,c,d)=s_i(a,b)+b_i(c,d) \, ,
\end{equation}
and is thus a function of four parameters. 
The four parameters $a,b,c,d$ have been introduced in order to describe
systematical uncertainties and are the nuisance parameters.
Each of the four parameters has a central value and systematical error.
The central values for all of
the four parameters have to be always defined. They are called 
signal normalization ($a$), signal tilt/calibration ($b$), 
background  normalization ($c$) and
background tilt/calibration ($d$). The default values are
\begin{equation}
a=1\,,\quad b=0\,,\quad c={\tt not \, \, assigned}\,,\quad d=0\,.
\end{equation}
Thus, for the background normalization $c$, the value has to be specified 
in \emph{either} case. The values for the normalization and the 
values of tilt/calibration are always regarded
as a pairs, \ie, they are given in the form {\tt normalization : tilt}. The errors are treated in the same way. For example, we have
\index{aedl}{{\tt rule}!signal error@{\tt "@signalerror}}
\index{aedl}{{\tt rule}!background error@{\tt "@backgrounderror}}
\index{aedl}{{\tt rule}!background center@{\tt "@backgroundcenter}}
\begin{quote}
{\tt
\tb @signalerror =       0.001  :       0.01\\
\tb @backgroundcenter =  0.1 :       0.0\\
\tb @backgrounderror =   0.001 :       0.01
}
\end{quote}
There is no {\tt @signalcenter} in this definition, 
since by default the central value for the
signal normalization is $1$ and the central value for the tilt/calibration 
is $0$.

The user has the possibility to choose the set $\{\zeta_i\}$ of nuisance 
parameters which are minimized over. This choice is specified with the 
error dimension variable\index{norm}{Error dimension}, and the different 
possibilities are shown in
\Tab~\ref{tab:error_dim}. Since the error dimension defines the treatement
of systematics it is useful to have define a matched pair of error dimensions
for each rule, where on value describes how the event rate is computed
for \emph{no} systematics and the other one is with systematics 
(see also~\Sec~\ref{sec:systematics}).
The error dimensions for the case of no systematics is set for each rule 
by the value of {\tt @errordim\_sys\_off}\index{aedl}{{\tt rule}!errordim 
sys off@{\tt "@errordim\_sys\_off}}, whereas the error dimension
for systematics on is given by {\tt @errordim\_sys\_on}
\index{aedl}{{\tt rule}!errordim sys on@{\tt "@errordim\_sys\_on}}. Thus,
 for example,
the complete error dimension definition could look like  
\begin{quote}
{\tt 
\tb @errordim\_sys\_off = 2\\
\tb @errordim\_sys\_on = 0
}
\end{quote}
It is foreseen to add the possiblity to extend the set 
of error dimensions or the set of possible systematical errors in one
of the next versions of \GLOBES.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{center}
\begin{table}[t!]
\begin{center}
\begin{tabular}[h]{|c|cccc|c|l|}
\hline
Error dimension&$a$&$b$&$c$&$d$&Tilt/Calibration&Remarks\\
\hline
\hline
0&+&+&+&+&T&Systematics with tilt\\
2&-&-&-&-&-&No systematics\\
4&+&+&+&+&T&Total rates\\
7& $\infty$ &-& $\infty$ &-&-&Spectrum only\\
8&-&-&-&-&-&Total rates, No systematics\\
9&+&+&+&+&C&Systematics with calibration\\
\hline
\end{tabular}
\caption[Table of error dimensions]{\label{tab:error_dim}
Possible values of the error dimensions variable in \GLOBES\ and their meaning. If a parameter is designated with $+$, it will
be marginalized over, and therefore the corresponding error needs to 
have a non-zero value. If the cases with ``total rates'' in the remarks, 
the summation over the bins is performed \emph{before} computing 
the $\chi^2$, \ie, no spectral information is used. 
The error dimension~7 ``spectrum only'' leaves the 
normalization free ($\sigma_a=\sigma_c=\infty$), and therefore 
only the spectral information is used. As a consequence,
the settings for the normalization error will be ignored (designated 
with the symbol $\infty$).

 }
\end{center}
\index{aedl}{{\tt rule}!errordim@{\tt "@errordim}} 
\end{table} 
\end{center}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

\AEDL\ is a powerful language to describe a vast variety 
of different experiments.
This chapter demonstrates how to test whether a given \AEDL\ file describes
a given experiment. To this end the \GLOBES\ package also features a program
called {\tt globes}\index{norm}{globes@{\tt globes}}, which can be seen either as a 
crude \AEDL\ debugger 
or a simple command line oriented tool to convert the rather 
abstract \AEDL\ experiment description into more accessible event rates.

@@

The {\tt globes}\index{norm}{globes@{\tt globes}} binary is installed along 
with the library \index{norm}{libglobes@{\tt libglobes}}, but
into the directory {\tt \$prefix/bin/}. In order to use the {\tt globes} 
utility this directory\footnote{This is automatically the case, when
no options have been given to {\tt configure}, and {\tt make install} was 
executed with root-privilege, \ie\ a standard installation was done.} has to 
be in the path of the shell used to invoke it.

@@

\index{norm}{globes@{\tt globes}!oscillation parameters}
Calling {\tt globes} without any options and with a {\tt .glb}-file as argument produces an event summary at rule level, taking into account 
the full description in the file, \ie\ all efficiencies, backgrounds and resolution effects. Thus, these are the event rates which are actually 
used to compute the $\chi^2$. By default the oscillation parameters used to
calculate the transition probability are
\begin{eqnarray}
\label{eq:globes_params}
\sin^22\theta_{12}=0.8&\quad&\Delta m^2_{21}=7\cdot10^{-5}
\,\mathrm{eV}^2\,,\nonumber\\
\sin^22\theta_{23}=1.0&\quad&\Delta m^2_{31}=3\cdot10^{-3}
\,\mathrm{eV}^2\,,\nonumber\\
\delta=0&\quad&\sin^22\theta_{13}=0.1\,.\
\end{eqnarray}
Of course, it is possible to change these default values either by using the
option {\tt -p} on a call by call basis or by setting the environment variable
{\tt GLB\_CENTRAL\_VALUES}
%
\index{norm}{GLB CENTRAL VALUES@{\tt GLB\_CENTRAL\_VALUES}} 
\index{norm}{Environment variables!GLB CENTRAL VALUES@{\tt GLB\_CENTRAL\_VALUES}}
%
\begin{quote}
{\tt
globes -p'0.55,0,0.785,0,0.0008,0.0025'\\
globes \verb+--+parameters='0.55,0,0.785,0,0.0008,0.0025'\\
}
\end{quote}
or by defining within the shell session or in the shell profile:
\begin{quote}
{\tt
export GLB\_CENTRAL\_VALUES='0.55,0,0.785,0,0.0008,0.0025'
}
\end{quote}
Furthermore it is possible to switch off oscillations altogether with t
he {\tt -N} option
and to switch them on again with {\tt -O} (the default). The effect of {\tt -N}
is the same than to prefix all channels with {\tt NOSC\_}.

\index{norm}{globes@{\tt globes}!warnings}
\index{norm}{globes@{\tt globes}!errors}
\index{norm}{globes@{\tt globes}!verbosity}
The \AEDL\ parser and interpreter has basically three levels of messages to
th user: warnings, errors and fatal errors. Fatal errors always are reported
and the program exits with status '1'. Usually only errors and no warnings
are reported. The verbosity level can be chosen by the {\tt -v} option, where
{\tt -v1} is default, \ie\ only errors and fatal errors are reported. {-v0}
corresponds to reporting fatal errors only, whereas {\tt -v2} will
print in addition to fatal errors and errors also warnings. It is recommended
to test any new {\tt .glb}-file with {\tt -v2} to see the warnings at least 
once and to decide whether there is a problem and in case fix it. With {\tt -v3}
all files read by {\tt globes} are displayed together with their path and
with  {\tt -v4} all files which have been attempted to be read are shown.
These two setting are useful to clarify path resolution issues and shadowing
of file names.

@@

\index{norm}{globes@{\tt globes}!total rates}
\index{norm}{globes@{\tt globes}!spectral rates}
By default {\tt globes} returns total rates corresponding to the 
{\tt -t} option, this can be changed to
to a full spectrum by using {\tt -s}. The spectral rates are shown in a
table where the first columns always gives the central energy of the
corresponding bin.

@@

\subsection*{Channel level}
\index{norm}{globes@{\tt globes}!channel rates}

@@

\subsection*{Rule level}
\index{norm}{globes@{\tt globes}!rule rates}

@@

\subsection*{Output}
\index{norm}{globes@{\tt globes}!output}

@@

\subsection*{\AEDL\ variable substitution}
\index{norm}{globes@{\tt globes}!variable substitution}
